---
title: "PUBH 7462 Week 2 Activity"
author: 'Jack Rabe'
date: "1/28/22"
output: 
  github_document:
    toc: true
    toc_depth: 1
  
---

```{r setup, include = FALSE}
#Load all the good stuff
library(tidyverse)

#Working directory for .RMD
knitr::opts_knit$set(echo = TRUE,
                     root.dir = rprojroot::find_rstudio_root_file())

#Controlling figure output in markdown
knitr::opts_chunk$set(
#  fig.height =   
  fig.width = 6,
#  fig.asp = .5,
  out.width = "90%",
#  out.height = 
 fig.align = "center",
  cache = FALSE,
  echo  = TRUE
)

#Set Themes for ggplot2 - centers title and legend at bottom by default and edit background
theme_set(theme(plot.title = element_text(hjust = 0.5),
                legend.position = "bottom", 
                panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                panel.background = element_blank(), 
                axis.line = element_line(colour = "black"))
)


#Set Scientific notation output and decimal places for knitr
options(scipen = 999)
options(digits = 4)
```

# Instructions  

1. Set up a new Github repo called `pubh_7462_week_2`
2. Clone it to your local machine in Rstudio w/ .Rproj
3. Edit the .gitignore (add /data line)
4. Download [Week 2 Activity RMD](https://canvas.umn.edu/courses/293049/assignments/2401738?module_item_id=7611944), save it in your directory, and edit the document to include the following:   
    - Change output to knit to [Github Document](https://rmarkdown.rstudio.com/github_document_format.html) (`output: github_document`)
    - Add a table of contents (`toc: true`)
    - Change the toc depth (`toc_depth: 1`)
    - Download `noaa_weather.RDS` file [here](), add a `/data` folder to your local directory, put this file there  
    - Add to the `ggplot` below as prompted in the .RMD (below)
    - Knit the document to `github_document` .md   
5. Use the Github workflow to push the .md and associated files to Github
6. Go back to Github, view your repository, and check out the .md  

\pagebreak

# National Oceanic and Atmospheric Association (NAOO) `ggplot` Example  

Here is a good example of data wrangling in the tidyverse, _however_ the `ggplot` visualization is missing a couple key elements (). Please fix the missing aesthetics that would make it a good, stand-alone data visualization.

## NAOO Data Wrangling  

This code chunk loads and tidies an example NAOO dataset, which extracts the 2017 daily record(s) of temperature min, max, and precipitation at each of three weather stations (NY, Hawaii, U of M). Notice that the chunk has the `eval = FALSE` option, this is because accessing the NOAA API takes a bit of time. __Please don't run this code in class__.

Instead, I've written out an .RDS file of the data we need and put it on Canvas [here](https://canvas.umn.edu/courses/293049/assignments/2401738?module_item_id=7611944). Please create a `/data` folder in the `pubh_7462_week_2` directory on your local machine and put the .RDS file there. We will load that data in the next chunk with a _relative path_.  

```{r noaa_tidy_example, echo = TRUE, eval = FALSE}
#Load RNOAA interface (and install if it's not already)
if (!require("rnoaa")) {
  install.packages("rnoaa")
}

#Create a weather data frame  
weather.df <- rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USC00214106"), #Location IDs
    var = c("PRCP", "TMIN", "TMAX"), #Variables 
    date_min = "2017-01-01", #Dates
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, 
           USW00094728 = "Central Park, NY", #Recode IDs variable 
           USC00519397 = "Waikiki, HA",
           USC00214106 = "U of M, MN"
           ),
    tmin = tmin / 10, #Scale temperature for viz I think it's in Kelvin? 
    tmax = tmax / 10) %>%
  select(name, id, everything())

#Write out as .RDS (generic R data file, i.e. a data frame, list, etc.)
write_rds(weather.df, "./data/roaa_weather.RDS")
```


\pagebreak 

## Example `ggplot` of Max Temperature  

```{r warning = FALSE, message = FALSE}
#Read in noaa_weather.RDS
weather.df <- read_rds("./data/roaa_weather.RDS")

#Create ggplot
max_temp.gg <- weather.df %>% 
  ggplot(aes(x = date, y = tmax, colour = name)) + 
  geom_point(alpha = 0.24, #Transparency
             size  = 2,
             shape = 16) +
  geom_smooth(formula = y ~ x, 
              method  = "loess", #Smooth mean trend
              se      = FALSE) #No Confidence interval, crowds plot

#Display max_temp.gg
max_temp.gg
```

\pagebreak 

## Your Turn  

### 1. Add missing aesthetics to the above Max Temp plot  

Create a new code chunk and please add these aesthetics to the ggplot above:  

- meaningful title  
- meaningful axis labels  
- colour pallete + meaningful legend title  
    - `+ scale_colour_viridis_d("Title")`  
- **Bonus**  
    - use `forcats::fct_reorder()` to reorder the id name by the maximum Max Temp in descending order (from highest to lowest) 
    - will have to transform to factor first w/ `as.factor()`

You may do so by simply writing `max_temp.gg + ...`  

```{r}
#Bonus
#reorder id name by the maximum max temp in descending order for plot
#first change id to factor
weather.df$id <- as.factor(weather.df$id) %>% 
  #now reorder accordingly
  forcats::fct_reorder(weather.df$id, weather.df$tmax, .fun = max, .desc = TRUE)

```


```{r warning = FALSE, message = FALSE}
#recreate plot with new aesthetics
max_temp.gg +
  labs(title = "NOAA Max Temperature across three locations in 2017",
       x = "Date",
       y = "Scaled Max Temperature") +
  scale_color_viridis_d("Location")
```


### 2. (If time allows) Create a `ggplot` of precipitation by location

```{r warning = FALSE, message = FALSE}
#plot of precipitation over time by location
ggplot(weather.df, aes(x = date, y = prcp, color = name)) +
  geom_point(alpha = 0.24, #Transparency
             size  = 2,
             shape = 16) +
  geom_smooth(formula = y ~ x, 
              method  = "loess", #Smooth mean trend
              se      = FALSE) + #No Confidence interval, crowds plot
  labs(title = "NOAA precipitation across three locations in 2017",
       x = "Date",
       y = "Precipitation (mm)") +
  scale_color_viridis_d("Location")
```


### 3. Push to Github  

If you've completed the first or both tasks, then please knit the document as a `github_document`. Next, please add, commit, and push your changes to the Github repository connected to this project. Note that you will need to push all the associated files (contains the .html for the figures you've created) as well as the .md and .rmd. In addition, you may need to create a personal access token ([with these instructions](https://docs.github.com/en/authentication/keeping-your-account-and-data-secure/creating-a-personal-access-token)) and use that as a password when prompted to login to github. Finally, go back to your github and check out your fine work by clicking on the newly knitted .md file in your repository!    

